The SMT Competition 2015--2018

### 2018-10-16 2018/report/SMT-COMP.tex
Divisions and Benchmarks (work in progress)

Benchmarks in the SMT Library are written in the SMT-LIB
language~\cite{BarFT-RR-17}, a text-based format that defines the
syntax and semantics of solver input and output.  This includes a
language for terms and formulas, and a command language for
interacting with SMT solvers.  For instance,
Figure~\ref{fig:smtlib-example} shows the example formula from
Section~\ref{sec:introduction} as a benchmark in SMT-LIB syntax.

@@

The SMT Library classifies benchmarks according to their logic, that
is, according to the specific combination of background theories that
the benchmark uses.  Logic names are composed of letter groups that
refer to these background theories: {\tt A} or {\tt AX} for the theory
of arrays, {\tt BV} for the theory of bit-vectors, {\tt DT} for
algebraic datatypes, {\tt FP} for the floating-point theory, {\tt IA}
for integer arithmetic, {\tt RA} for real arithmetic, {\tt IRA} for
mixed integer and real arithmetic, {\tt S} for the theory of strings,
{\tt UF} for uninterpreted functions (i.e., free symbols).
Additionally, logics may impose further syntactic restrictions, such
as the absence of quantifiers ({\tt QF}), the restriction to
difference logic over the integers ({\tt IDL}) or reals ({\tt RDL}),
or to the (non-)linear fragment of arithmetic ({\tt L} or {\tt N},
respectively, before {\tt IA}, {\tt RA} or {\tt IRA}).  For instance,
the logic {\tt QF\_UFLIA} contains quantifier-free formulas with
uninterpreted functions over linear integer arithmetic.  These
restrictions are typically motivated by the existence of efficient
decision procedures for certain syntactic fragments.

Additionally, the SMT Library distinguishes between incremental and
non-incremental benchmarks.  Non-incremental benchmarks contain a
single {\tt check-sat} command, which instructs the SMT solver to
determine the satisfiability of the benchmark.  (Valid solver
responses are {\tt sat}, {\tt unsat}, and {\tt unknown}.)  Incremental
benchmarks exercise additional solver features, such as the ability to
retract assertions, and typically contain multiple {\tt check-sat}
commands.  These benchmarks originate from applications such as
bounded model checking~\cite{TODO}, where SMT solvers interact with
other tools in a feedback loop.  The 2018 release of the SMT Library
contained 336,844 non-incremental benchmarks in 51 logics, as well
as~10,167 incremental benchmarks (with a total of nearly 32 million
{\tt check-sat} commands) in 26 logics.  Table~\ref{table:smtlib}
presents further data on the size of the SMT Library since~2015.
While the library usually grows from one year to the next due to the
inclusion of new benchmarks and logics, continuous curation efforts
may also cause benchmarks to be removed or reclassified on occasion,
thereby causing individual logics or even the entire library to
shrink.

@@

>>>>

Benchmarks in the SMT Library are written in the SMT-LIB
language~\cite{BarFT-RR-17}, a text-based format that defines the
syntax and semantics of solver input and output.  This includes a
language for terms and formulas, and a command language for
interacting with SMT solvers.  For instance,
Figure~\ref{fig:smtlib-example} shows the example formula from
Section~\ref{sec:introduction} as a benchmark in SMT-LIB syntax.
In~2016, benchmarks were updated from version~2.0 to version~2.5 of
the SMT-LIB language, and since~2017, benchmarks are written in
version~2.6.  Fortunately, versions~2.5 and~2.6 are largely
backwards-compatible to 2.0, so that older solvers are still able to
compete with at most minor modifications.

@@

The SMT Library classifies benchmarks according to their \emph{logic},
that is, according to the specific combination of background theories
that the benchmark uses.  Logic names are composed of letter groups
that refer to these background theories: {\tt A} or {\tt AX} for the
theory of arrays, {\tt BV} for the theory of bit-vectors, {\tt DT} for
algebraic datatypes, {\tt FP} for the floating-point theory, {\tt IA}
for integer arithmetic, {\tt RA} for real arithmetic, {\tt IRA} for
mixed integer and real arithmetic, {\tt S} for the theory of strings,
{\tt UF} for uninterpreted functions (i.e., free symbols).
Additionally, logics may impose further syntactic restrictions, such
as the absence of quantifiers ({\tt QF}), the restriction to
difference logic over the integers ({\tt IDL}) or reals ({\tt RDL}),
or to the (non-)linear fragment of arithmetic ({\tt L} or {\tt N},
respectively, before {\tt IA}, {\tt RA} or {\tt IRA}).  For instance,
the logic {\tt QF\_UFLIA} contains quantifier-free formulas with
uninterpreted functions over linear integer arithmetic.  These
restrictions are typically motivated by the existence of efficient
decision procedures for certain syntactic fragments.

Additionally, the SMT Library distinguishes between \emph{incremental}
and \emph{non-incremental} benchmarks.  Non-incremental benchmarks
contain a single {\tt check-sat} command, which instructs the SMT
solver to determine the satisfiability of the benchmark.  (Valid
solver responses are {\tt sat}, {\tt unsat}, and {\tt unknown}.)
Incremental benchmarks exercise additional solver features, such as
the ability to retract assertions, and typically contain multiple {\tt
  check-sat} commands.  These benchmarks originate from applications
such as bounded model checking~\cite{TODO}, where SMT solvers interact
with other tools in a feedback loop.  The 2018 release of the SMT
Library contained 336,844 non-incremental benchmarks in 51 logics, as
well as~10,167 incremental benchmarks (with a total of nearly 32
million {\tt check-sat} commands) in 26 logics.
Table~\ref{table:smtlib} presents further data on the size of the SMT
Library since~2015.  While the library usually grows from one year to
the next due to the inclusion of new benchmarks and logics, continuous
curation efforts may also cause benchmarks to be removed or
reclassified on occasion, thereby causing individual logics or even
the entire library to shrink.

@@

The SMT Competition consists of several tracks.  The \emph{main
  track}, which has been a staple of the competition since its
inception in 2005, features non-incremental benchmarks.  The
\emph{application track}, which has been part of the competition since
2011~\cite{TODO}, features incremental benchmarks.  These benchmarks
are passed to each solver one command at a time by a trace executor,
which waits for the solver's response before passing the next
benchmark command (to prevent look-ahead behavior).  The
\emph{unsat-core track} uses main-track benchmarks that are known to
be unsatisfiable, and solvers are evaluated on their ability to find
small unsatisfiable cores (i.e., a small but unsatisfiable subset of
the benchmark's assertions).  The unsat-core track had first been part
of the competition in 2012~\cite{TODO}, but was then discontinued
partly because of the competition's move to StarExec (see Section~5)
and lacking tool support.  It was re-introduced as an experimental
track in 2016, and has been a regular competition track since~2017.

Each track is further split into (logic) \emph{divisions}, where each
division corresponds to a logic from the SMT Library.  Participants
can choose which tracks and divisions they want to enter, and
competition results are reported separately for each track and
division.  Thus, competing solvers need not support all logics, and
may or may not support features such as incrementality or unsat-core
generation.

SMT Competitions until~2014 used a pseudo-random selection process to
cull the number of benchmarks used in the competition.  However, a
2013 evaluation found that this ``significantly lessens the ability of
a competition to determine `best'
solvers.''~\cite{DBLP:journals/jar/CokSW15} Together with the
increased computational resources available for the competition since
its move to StarExec, this prompted a paradigm shift.  Since 2015, the
competition is evaluating all solvers on all eligible benchmarks.
Some benchmarks from the SMT Library typically have to be excluded for
syntactic reasons: in 2015 and 2016, 11,966 benchmarks were excluded
because they contained partial or underspecified operations (e.g.,
bit-vector division) whose precise semantics were being debated at the
time.  Additionally in 2016, one benchmark was excluded because it
contained incorrect status information.  In 2017, 1,104 benchmarks
were excluded because their logic had been incorrectly classified, and
8 benchmarks were excluded because of incorrect or missing status
information.  In 2018, 3,603 benchmarks were excluded because they
contained named terms (a feature of the SMT-LIB language that is not
permitted in competition benchmarks), and 5 benchmarks were excluded
because of missing status information.

\begin{tjark}
  WIP: benchmark selection, tables with numbers, excluded benchmarks,
  unknown benchmarks
\end{tjark}
